Simultaneous Elicitation of Scoring Rule and Agent Preferences for Robust Winner Determination

### 2021-01-20 minimax.tex
take away back to intro

Our approach is evaluated on simulations with synthetic and real datasets where both the voting rule and the agent preferences are initially unknown to the system and incrementally revealed through questioning.
We compare the effectiveness of several questioning strategies based on the current knowledge of the rule and preferences. 
\section{Social choice with partial information}
\label{sec:background}
We now introduce some basic concepts.
We consider a set $A$ of $m$ alternatives (products, restaurants, public projects, job candidates, etc.), an infinite set $\N$ of potential agents and a specific finite set of agents $N^* \subset \N$.

@@

%\begin{figure}
%	\caption{Average MMR in problems of size $(5, 10)$ after $k$ questions.}
%	\label{fig:smallSize}
%%	\Description{Graph showing that Pessimistic performs slightly better than Extended pessimistic, itself performing better than Elitist, itself performing much better than Random.}
%	\includegraphics[width=.45\textwidth]{comparison.png}
%\end{figure}
\sisetup{table-number-alignment = center, table-figures-integer=2, table-figures-decimal=1, table-auto-round}
\begin{table}
	\caption{Average MMR in problems of size $(10, 20)$ after $k$ questions assuming geometric weights.}
	\label{tab:geometricWeights}
	\begin{tabular}{S[table-figures-integer=3, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left, table-figures-integer=1]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left, table-figures-integer=1]}
		\toprule
		{k} & {Pes.} & {sd} & {Eli.} & {sd} \\
		\midrule
		0	&20.0	&0.0	&20.0	&0.0\\
		50	&15.96	&0.54	&17.26	&0.42\\
		100	&12.48	&0.93	&15.6	&0.43\\
		150	&9.58	&1.37	&13.94	&0.76\\
		200	&7.43	&1.25	&10.95	&1.12\\
		250	&5.26	&1.52	&6.6	&0.79\\
		300	&3.47	&1.32	&6.57	&0.79\\
		\bottomrule
	\end{tabular}
\end{table}
\paragraph{Comparison of strategies}
Our first experiment concerns small size situations.
\Cref{fig:smallSize} compares some of our strategies in the case $m = 5, n = 10$ (variations around this size yield similar conclusions), where the results are averaged over $200$ runs.
%(The figure also displays the performance of the Elitist strategy, to which we will come back.)
We see that asking random questions does not allow to reach a low regret level even after having asked 100 questions, whereas a low regret level ($\MMR$ = 1) is reached by Pessimistic before having asked 60 questions.
We also see that Pessimistic performs slightly better than Extended pessimistic, showing that Pessimistic chooses candidate questions wisely; this is good news since Pessimistic is much faster: it takes on average only $16$s for a complete elicitation session (for $m = 5$, $n = 10$ and $100$ questions), while Extended pessimistic takes $50$s. Although their performance is close, Pessimistic performs systematically better in multiple runs of the experiment.
\Cref{sec:appendix} provides a problem of larger size.
\commentPV{Check problem size displayed in the labels of the two table. The one in the appendix does not seem larger.}

@@

We summarize the main observations from the experiments:
\commentPV{We could save space by 'de-itemizing' this, i.e. by formatting as text with i) bla bla bla.... ii) .... iiii) ....}
\begin{itemize}
%	\item we provide a novel mechanism for eliciting a voting rule by translating abstract questions about weights to a choice of an alternative given a concrete profile;  
	\item we showed that with our elicitation method %in small problem sizes, %starting with zero knowledge,
	 it is possible to reach low regret with a reasonable number of questions;
	\item the elicitation strategies that we provide, in particular Pessimistic, achieve good results within reasonable computation time; %for small problem sizes;
	\item we showed that for the class of rule considered, asking a few questions to the chair suffice to reach  low regret;
	\item our experiments suggest that low degree of similarity among preferences %rankings 
	(as in impartial culture) is a more challenging setting than less varied profiles.
\end{itemize}

@@

\newpage
\appendix
\section{Appendix}
\label{sec:appendix}
\newtheorem*{details*}{Strategy Details}

\begin{details*}[Pessimistic]
	Given $j \in N^*$, if $x^*$ and $\bar{y}$ are incomparable in $\ppref_j$, the candidate question concerns the pair $(x^*, \bar{y})$, otherwise, the candidate question concerns the pair $(x^*, z)$ for some $z$ incomparable to $x^*$ (randomly chosen), or if none such $z$ exist, the pair $(\bar{y}, z)$ for some $z$ incomparable to $\bar{y}$, or, if both $x^*$ and $\bar{y}$ are comparable to every alternatives in $\ppref_j$, any incomparable pair is picked at random. 
%Responses to questions involving $x^*$ or $\bar{y}$ constrain the ability of the adversary to complete the profile in such manner, possibly reducing minimax regret.
	Once having selected $n + m$ candidate questions, the Pessimistic strategy selects the one that leads to minimal regret in the worst case, considering both possible answers to the question, and with penalty terms depending on the kind of question. To define this comparison precisely, assume that a question $q_1$ has type $t_1$ (being “chair” or “agent”), and leads to the new knowledge states $(\pprofile_1, W_1)$ if answered positively and $(\pprofile'_1, W'_1)$ if the answer is negative. 
	Define $R^{\max}_1 = \max\set{\MMR(\pprofile_1, W_1), \MMR(\pprofile'_1, W'_1)}$
	and $R^{\min}_1 = \min\set{\MMR(\pprofile_1, W_1), \MMR(\pprofile'_1, W'_1)} \epsilon_{t} + \epsilon'_{t}$.
	%If $(\pprofile_1, W_1) ≥ (\pprofile'_1, W'_1)$, define $\MMR^{\max} = \MMR(\pprofile_1, W_1)$ and $\MMR^{\min} = \MMR(\pprofile'_1, W'_1)$; otherwise, define $\MMR^{\max} = \MMR(\pprofile'_1, W'_1)$ and $\MMR^{\min} = \MMR(\pprofile_1, W_1)$.
	% and and $(\pprofile^\min_2, W^\min_2)$. numbering them so that $\MMR(\pprofile_1, W_1) ≥ \MMR(\pprofile_2, W_2)$. Then the badness of the question in the worst case is:
	The terms $\epsilon_t$ and $\epsilon'_{t}$ are real numbers associated to the type $t$ of question; these parameters are to be determined experimentally and are used to fine control the choice of the question type.
	Define similarly $t_2$, $R^{\max}_2$ and $R^{\min}_2$ for a question $q_2$.
	We are now faced with the problem of aggregating these values in order to compare the informative values of the candidate questions. We adopt a pessimistic approach, reported to perform better than optimistic aggregation \citep{Cailloux2014}, although our experiments in this context suggested a weak impact of that choice on the performance of the strategy.
	To avoid the absorption property of the max, we adopt $\leximax$ as an aggregator.
	Pessimistic thus considers question $q_1$ to be better  than $q_2$ iff $R^{\max}_1 < R^{\max}_2 \text{ or } [R^{\max}_1 = R^{\max}_2 \text{ and } R^{\min}_1 < R^{\min}_2]$.
	%This comparison gives a way of picking questions among a set of possible questions, by picking one that minimizes this approximate measure of minimax regret {\em a posteriori}. 
	%The motivation for using this operator lies in 
	In other words if the maximal {\em a posteriori} MMR of two questions are (nearly) equal, then it considers the (penalized) minimal MMR values, preferring the question with the lowest value.
	(Technically, instead of applying a pure lexicographic aggregation, which is very sensitive to small errors due to floating-point computations, we apply a weighted sum between the maximal and the penalized minimal MMR values, with a predominant weight given to the maximal one, several order of magnitudes higher than the weight given to the penalized minimal MMR.)

>>>>

Our approach is evaluated on simulations with synthetic and real datasets where both the voting rule and the agent preferences are initially unknown to the system and incrementally revealed through questioning.
We compare the effectiveness of several questioning strategies based on the current knowledge of the rule and preferences.
\commentBN{on small sizes because we assume humans}
\commentBN{remove bullets and restore PMR}
\begin{itemize}
	\item we provide a novel mechanism for eliciting a voting rule by translating abstract questions about weights to a choice of an alternative given a concrete profile;  
	\item we showed that with our elicitation method %in small problem sizes, %starting with zero knowledge,
	it is possible to reach low regret with a reasonable number of questions;
	\item the elicitation strategies that we provide, in particular Pessimistic, achieve good results within reasonable computation time; %for small problem sizes;
	\item we showed that for the class of rule considered, asking a few questions to the chair suffice to reach low regret;
	\item our experiments suggest that low degree of similarity among preferences %rankings 
	(as in impartial culture) is a more challenging setting than less varied profiles.
\end{itemize}

\section{Social choice with partial information}
\label{sec:background}
We now introduce some basic concepts.
We consider a set $A$ of $m$ alternatives (products, restaurants, public projects, job candidates, etc.), an infinite set $\N$ of potential agents and a specific finite set of agents $N^* \subset \N$.

@@

%\begin{figure}
%	\caption{Average MMR in problems of size $(5, 10)$ after $k$ questions.}
%	\label{fig:smallSize}
%%	\Description{Graph showing that Pessimistic performs slightly better than Extended pessimistic, itself performing better than Elitist, itself performing much better than Random.}
%	\includegraphics[width=.45\textwidth]{comparison.png}
%\end{figure}
\sisetup{table-number-alignment = center, table-figures-integer=2, table-figures-decimal=1, table-auto-round}
\begin{table}
	\caption{Average MMR in problems of size $(10, 20)$ after $k$ questions assuming geometric weights.}
	\label{tab:geometricWeights}
	\begin{tabular}{S[table-figures-integer=3, table-figures-decimal=0]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left, table-figures-integer=1]S[table-number-alignment = right]@{ ± }S[table-number-alignment = left, table-figures-integer=1]}
		\toprule
		{k} & {Pes.} & {sd} & {Eli.} & {sd} \\
		\midrule
		0	&20.0	&0.0	&20.0	&0.0\\
		50	&15.96	&0.54	&17.26	&0.42\\
		100	&12.48	&0.93	&15.6	&0.43\\
		150	&9.58	&1.37	&13.94	&0.76\\
		200	&7.43	&1.25	&10.95	&1.12\\
		250	&5.26	&1.52	&6.6	&0.79\\
		300	&3.47	&1.32	&6.57	&0.79\\
		\bottomrule
	\end{tabular}
\end{table}
\paragraph{Comparison of strategies}
Our first experiment concerns small size situations.
\Cref{fig:smallSize} compares some of our strategies in the case $m = 5, n = 10$ (variations around this size yield similar conclusions), where the results are averaged over $200$ runs.
%(The figure also displays the performance of the Elitist strategy, to which we will come back.)
We see that asking random questions does not allow to reach a low regret level even after having asked 100 questions, whereas a low regret level ($\MMR$ = 1) is reached by Pessimistic before having asked 60 questions.
We also see that Pessimistic performs slightly better than Extended pessimistic, showing that Pessimistic chooses candidate questions wisely; this is good news since Pessimistic is much faster: it takes on average only $16$s for a complete elicitation session (for $m = 5$, $n = 10$ and $100$ questions), while Extended pessimistic takes $50$s. Although their performance is close, Pessimistic performs systematically better in multiple runs of the experiment.
\Cref{sec:appendix} provides a problem of larger size.

@@

@@

\newpage
\appendix
\section{Appendix}
\label{sec:appendix}
\newtheorem*{details*}{Strategy Details}
\commentBN{delete all this}
\begin{details*}[Pessimistic]
	Given $j \in N^*$, if $x^*$ and $\bar{y}$ are incomparable in $\ppref_j$, the candidate question concerns the pair $(x^*, \bar{y})$, otherwise, the candidate question concerns the pair $(x^*, z)$ for some $z$ incomparable to $x^*$ (randomly chosen), or if none such $z$ exist, the pair $(\bar{y}, z)$ for some $z$ incomparable to $\bar{y}$, or, if both $x^*$ and $\bar{y}$ are comparable to every alternatives in $\ppref_j$, any incomparable pair is picked at random. 
%Responses to questions involving $x^*$ or $\bar{y}$ constrain the ability of the adversary to complete the profile in such manner, possibly reducing minimax regret.
	Once having selected $n + m$ candidate questions, the Pessimistic strategy selects the one that leads to minimal regret in the worst case, considering both possible answers to the question, and with penalty terms depending on the kind of question. To define this comparison precisely, assume that a question $q_1$ has type $t_1$ (being “chair” or “agent”), and leads to the new knowledge states $(\pprofile_1, W_1)$ if answered positively and $(\pprofile'_1, W'_1)$ if the answer is negative. 
	Define $R^{\max}_1 = \max\set{\MMR(\pprofile_1, W_1), \MMR(\pprofile'_1, W'_1)}$
	and $R^{\min}_1 = \min\set{\MMR(\pprofile_1, W_1), \MMR(\pprofile'_1, W'_1)} \epsilon_{t} + \epsilon'_{t}$.
	%If $(\pprofile_1, W_1) ≥ (\pprofile'_1, W'_1)$, define $\MMR^{\max} = \MMR(\pprofile_1, W_1)$ and $\MMR^{\min} = \MMR(\pprofile'_1, W'_1)$; otherwise, define $\MMR^{\max} = \MMR(\pprofile'_1, W'_1)$ and $\MMR^{\min} = \MMR(\pprofile_1, W_1)$.
	% and and $(\pprofile^\min_2, W^\min_2)$. numbering them so that $\MMR(\pprofile_1, W_1) ≥ \MMR(\pprofile_2, W_2)$. Then the badness of the question in the worst case is:
	The terms $\epsilon_t$ and $\epsilon'_{t}$ are real numbers associated to the type $t$ of question; these parameters are to be determined experimentally and are used to fine control the choice of the question type.
	Define similarly $t_2$, $R^{\max}_2$ and $R^{\min}_2$ for a question $q_2$.
	We are now faced with the problem of aggregating these values in order to compare the informative values of the candidate questions. We adopt a pessimistic approach, reported to perform better than optimistic aggregation \citep{Cailloux2014}, although our experiments in this context suggested a weak impact of that choice on the performance of the strategy.
	To avoid the absorption property of the max, we adopt $\leximax$ as an aggregator.
	Pessimistic thus considers question $q_1$ to be better  than $q_2$ iff $R^{\max}_1 < R^{\max}_2 \text{ or } [R^{\max}_1 = R^{\max}_2 \text{ and } R^{\min}_1 < R^{\min}_2]$.
	%This comparison gives a way of picking questions among a set of possible questions, by picking one that minimizes this approximate measure of minimax regret {\em a posteriori}. 
	%The motivation for using this operator lies in 
	In other words if the maximal {\em a posteriori} MMR of two questions are (nearly) equal, then it considers the (penalized) minimal MMR values, preferring the question with the lowest value.
	(Technically, instead of applying a pure lexicographic aggregation, which is very sensitive to small errors due to floating-point computations, we apply a weighted sum between the maximal and the penalized minimal MMR values, with a predominant weight given to the maximal one, several order of magnitudes higher than the weight given to the penalized minimal MMR.)
